Bayesian Sparsification Methods for Deep Complex-valued Networks

### 2020-03-27 paper/paper.tex
cited and added Louizos et al. (2017) for hierarchical priors and group sparsity

The key role in the factorization is played by $\alpha_j$, which reflects the relevance
of the parameter $\omega_j$ it is associated to, by being the ratio of its squared mean
to its variance. Thus, \citet{molchanov_variational_2017} focus on the model sparsification
capabilities of Variational Dropout, optimizing $\alpha_j$ for each individual parameter,
or across groups of parameters for structured sparsity. \citet{kharitonov_variational_2018}
address theoretical issues with improper priors and propose a method based on \textit{
Automatic Relevance Determination} procedure, which falls under \textit{Empirical Bayes}
approach that fits a prior distribution while performing Variational Inference. In this
method the log-uniform factor $\pi(\omega_j)$ is replaced by a proper Gaussian ARD
prior $
  \mathcal{N}(\omega_j \vert\, 0, \tau^{-1}_j)
$ with learnable precision $\tau_j > 0$, \citep{neal_bayesian_1996}. Maximizing
\eqref{eq:elbo} over $\tau$, holding other parameters fixed, yields $
  \tau^*_j = {(\mu_j^2 + \sigma^2_j)}^{-1}
$, whence
\begin{equation}  \label{eq:ard-kl-div-real}
  KL(q_\theta \| \pi)
    = \frac12 \sum_j \log{\bigl(1 + \tfrac1{\alpha_j} \bigr)}
    \,.
\end{equation}

@@

More recently general development of $\cplx$-valued networks has been continued by
\citet{trabelsi_deep_2017}, who outline the essential building blocks for deep $\cplx$-valued
networks and describe suitable representation and operations, including convolutional and
dense layers, $\cplx$-valued activations, complex batch-normalization and weight initialization.
Their approach coupled with $\cplx\real$ calculus for backprop transforms a $\cplx$-valued
network into an intricately connected $\real$-valued computational graph, constructured
to respects $\cplx$-arithmetic.
%
In particular, natural identification of $\cplx$ and $\real^2$ implies that dense linear
layers $
  L \colon \cplx^m \to \cplx^n
$ act upon their inputs as follows:
\begin{equation}  \label{eq:cplx-lin-op}
  {(\real^n)}^{2}
    \to {(\real^m)}^{2}
    \colon (u, v)
      \mapsto \Bigl(
        P u - Q v,
        P v + Q u
      \Bigr)
    \,,
\end{equation}
where $
  P, Q \colon \real^{n} \to \real^{m}
$ are unique operators such that $L = P + j Q$. A $\cplx$-convolutional layer with kernel
$W$ can be implemented as two $\real$-convolutions with kernels $\Re{W}$ and $\Im{W}$ using
\eqref{eq:cplx-lin-op}. Activations include trigonometric functions and representation-dependent
non-linearities \citep{hirose_complex-valued_2009}: polar (phase-amplitude) $
  z \mapsto \sigma(\lvert z \rvert) e^{j \arg z}
$ or planar (real-imaginary) $
  z \mapsto \sigma(\Re z) + j \sigma(\Im z)
$ for some $\real$-valued non-linearity $\sigma$, possibly with learnable parameters
\citep{trabelsi_deep_2017,wolter_complex_2018}.

>>>>

The key role in the factorization is played by $\alpha_j$, which reflects the relevance
of the parameter $\omega_j$ it is associated to, by being the ratio of its squared mean
to its variance. Thus, \citet{molchanov_variational_2017} focus on the model sparsification
capabilities of Variational Dropout, optimizing $\alpha_j$ for each individual parameter,
or across groups of parameters for structured sparsity.
%
\citet{louizos_bayesian_2017} consider hierarchical prior and approximation, which couples
the parameters $\omega_j$ within a group $J$ through shared scaling $z_J$ and enables
pruning entire input features in each layer.
% {louizos_bayesian_2017} eq. (11) variance -> gaussian entropy -> informative bits 
%
\citet{kharitonov_variational_2018}
address theoretical issues with improper priors and propose a method based on \textit{
Automatic Relevance Determination} procedure, which falls under \textit{Empirical Bayes}
approach that fits a prior distribution while performing Variational Inference. In this
method the log-uniform factor $\pi(\omega_j)$ is replaced by a proper Gaussian ARD
prior $
  \mathcal{N}(\omega_j \vert\, 0, \tau^{-1}_j)
$ with learnable precision $\tau_j > 0$, \citep{neal_bayesian_1996}. Maximizing
\eqref{eq:elbo} over $\tau$, holding other parameters fixed, yields $
  \tau^*_j = {(\mu_j^2 + \sigma^2_j)}^{-1}
$, whence
\begin{equation}  \label{eq:ard-kl-div-real}
  KL(q_\theta \| \pi)
    = \frac12 \sum_j \log{\bigl(1 + \tfrac1{\alpha_j} \bigr)}
    \,.
\end{equation}
% \begin{align}
%   \pi(\omega)
%     &
%     = \prod_J \pi(s_J)
%       \prod_{j\in J} \pi(\omega_j \vert s_J)
%     \propto \prod_J \frac1{\lvert s_J \rvert}
%       \prod_{j\in J} \mathcal{N}(\omega_j \vert 0, s_J^2)
%     \,, \\
%   q_\theta(\omega)
%     &
%     % = \prod_J q_\theta(s_J)
%     %   \prod_{j\in J} q_\theta(\omega_j \vert s_J)
%     = \prod_J \mathcal{N}(s_J \vert \mu_J, \alpha_J \mu_J^2)
%       \prod_{j\in J} \mathcal{N}(\omega_j \vert \mu_j s_J, \sigma_j^2 s_J^2)
%     \,.
% \end{align}

@@

More recently general development of $\cplx$-valued networks has been continued by
\citet{trabelsi_deep_2017}, who outline the essential building blocks for deep $\cplx$-valued
networks and describe suitable representation and operations, including convolutional and
dense layers, $\cplx$-valued activations, complex batch-normalization and weight initialization.
Their approach coupled with $\cplx\real$ calculus for backprop transforms a $\cplx$-valued
network into an intricately connected $\real$-valued computational graph, constructed
to respects $\cplx$-arithmetic.
%
In particular, natural identification of $\cplx$ and $\real^2$ implies that dense linear
layers $
  L \colon \cplx^m \to \cplx^n
$ act upon their inputs as follows:
\begin{equation}  \label{eq:cplx-lin-op}
  {(\real^n)}^{2}
    \to {(\real^m)}^{2}
    \colon (u, v)
      \mapsto \Bigl(
        P u - Q v,
        P v + Q u
      \Bigr)
    \,,
\end{equation}
where $
  P, Q \colon \real^{n} \to \real^{m}
$ are unique operators such that $L = P + j Q$. A $\cplx$-convolutional layer with kernel
$W$ can be implemented as two $\real$-convolutions with kernels $\Re{W}$ and $\Im{W}$ using
\eqref{eq:cplx-lin-op}. Activations include trigonometric functions and representation-dependent
non-linearities \citep{hirose_complex-valued_2009}: polar (phase-amplitude) $
  z \mapsto \sigma(\lvert z \rvert) e^{j \arg z}
$ or planar (real-imaginary) $
  z \mapsto \sigma(\Re z) + j \sigma(\Im z)
$ for some $\real$-valued non-linearity $\sigma$, possibly with learnable parameters
\citep{trabelsi_deep_2017,wolter_complex_2018}.
